---
title: "Biol 3872 Lab7-Assignment"
output: html_document
date: "`r Sys.Date()`"
author: ""
student_id: ""
---

\
**BEFORE STARTING:**\
1. Include your name (as shown in Brightspace) in the `author` section above as text (e.g., "Liam MacNeil") and Banner ID ("B00123456") in `student_id`.\
2. Install and/or load the packages required to run your code chunks.

Assignments will largely be evaluated based on your code chunk answers. If they are incomplete or formatted incorrectly (e.g., code chunks don't run), no marks will be awarded. I recommend experimenting with solving these answers in a separate script/markdown document and reporting your answers with minimal script.

```{r}
library(tidyverse)
library(terra)
library(sf)
library(rnaturalearth)
library(sdmpredictors)
```

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 1 {3 points}

A)  We have created a mock raster (`rast`) below, try plotting it. Store the summary of this raster layer in a variable called `rast_summary`

B)  Store it's spatial extent in a variable called `rast_extent`

C)  Store it's spatial resolution in a variable called `rast_res`.

```{r}
rast <- rast(ncol = 50, nrow = 50,  xmin = 1, xmax = 1.2, ymin = 1, ymax = 1.2)
values(rast) <- sample(1:ncell(rast), ncell(rast))

# YOUR CODE HERE

# A)

  
# B)


# C)


```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 2 {3 points}

A)  We have generated another mock raster (`rast_gauss`) which computes a gaussian distribution of values across the same grid. Stack this raster with `rast` from Q1 and call this `rast_stack`.

B)  Use `names` to modify the layer names of `rast_stack` to `Random` (`rast`) and `Gaussian` (`rast_gauss`).

C)  Average these two rasters and store it in a raster called `mean_rast`.

```{r}
rast_gauss <- rast(ncol = 50, nrow = 50,  xmin = 1, xmax = 1.2, ymin = 1, ymax = 1.2)
coords <- crds(rast_gauss, df = TRUE)
# Define middle of distribution
x0 <- (ext(rast_gauss)[1] + ext(rast_gauss)[2]) / 2 
y0 <- (ext(rast_gauss)[3] + ext(rast_gauss)[4]) / 2  
# Gaussian params
sigma <- 10  
# Calculate Gaussian values and assign
values(rast_gauss) <- exp(-((coords$x - x0)^4 + (coords$y - y0)^2) / (2 * sigma^2))

# YOUR CODE HERE
# A)


# B) 


# C)


```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 3 {3 points}

A)  Read in the Ecoregions polygons from your `data` folder to an `sf` object called `Ecoregions`.

B)  Filter those Ecoregions for polygons in the "Tropical Atlantic" `REALM`, and project the CRS for `trop_atl` to `EPSG:3857` (Mercator WGS84 datum, default of google maps). Call this new `sf` object `trop_atl`.

C)  Plot `trop_atl` using `ggplot`, remove the legend, and fill the Ecoregions by name. Call this `ggplot` object `trop_atl_p.`

```{r}
# YOUR CODE HERE
# A)


# B)


# C)

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 4 {2 points}

Load in the coastline file from `rnaturalearth` at "medium" resolution. Name this `coastline`.

A)  Crop this coastline file to the extent (bounding box) of the `trop_atl` Ecoregions. Call this `trop_coasts`. Are the CRS identical between shapefiles?

B)  Plot `trop_coasts` and `trop_atl` together using `ggplot` (which order should they be in?), filling ecoregions by name, and filling coastline with `grey70.` Remove the legend. Call this `ggplot` object `trop_atl_coasts_p`.

```{r}
# YOUR CODE HERE
# A)

# B)

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 5 {4 points}

A)  Read in `Groundfish_surveys.csv` from your data folder.

B)  Convert this data frame to an `sf` object called `trawls_sf` and filter for only rows containing species (`accepted_name`) Silver Hake (`Merluccius bilinearis`). Assign the CRS for `trawls_sf` to `EPSG:4326` (WGS84 datum).

C)  Use `trawls_sf` to index `Ecoregions` which intersect points (Do not crop by extent, but keep full polygon if they intersect anywhere). Then use `trawls_sf` to crop the global `coastline` shapefile (are these already in a compatible CRS?). Name these `Ecoregions_hake` and `coastline_hake`, respectively.

D)  Plot `trawls_sf` points, `Ecoregions_hake`, and `coastline_hake` using `ggplot` (which order should they be in?). Make `Ecoregions_hake` transparent (`fill=?`), fill `coastline_hake` with `grey20`, color `trawls_sf` points `purple` with 25% transparency, and add a `minimal` theme. Name this plot `NW_trawls_p`.

```{r}
# YOUR CODE HERE
# A)


# B)


# C)


# D)

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 6 {3 points}

A)  Load and stack the following layers from `sdmpredictors`: `BO_bathymean` (mean bottom depth), `BO_sstmean` (mean sea surface temp), `BO22_tempmax_bdmax` (max temp at bottom), `BO2_curvelmean_ss` (mean current velocity). Call this `rast_stack`.

B)  Spatially mask and crop `rast_stack` to the exact shape of `Ecoregions_hake`. Call this `rast_stack_hake`.

C)  Write a simple `for()` loop which, across layer `i` in `rast_stack_hake` (`1:nlyr(raster)`), plots the natural `log` transform of the raster layer. If you have indexed the loop properly, four plots should be produced. Assign a nice color palette of your choice– make the number of colors small (\~10).

```{r}
# YOUR CODE HERE
# A)


# B)


# C)

```
:::

\
**BEFORE SUBMITTING**

Clear your environment (broom) and run your entire script to ensure everything has been defined properly!

Be sure to submit this `markdown` file to the relevant Brightspace submission box.
